---
title: "p8105_hw3_bac2214"
author: "Brianna Carnagie"
date: "2023-10-08"
output: github_document
---
### Question 1
```{r,echo=FALSE, message=FALSE}
library(p8105.datasets)
library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```
#### Loading the data
```{r}
data("instacart") 
```
#### Exploring the data
```{r}
instacart_reordered = 
  instacart |> 
  select(order_id, user_id, order_number, everything())

user_112108 = 
  instacart |>  
  filter(user_id == 112108)

user_17989 = 
  instacart |>  
  filter(user_id == 17989) 

user_17989_dept_frq = 
  user_17989 |> 
  count(department_id)
```

The dataset is a tibble that has `r nrow(instacart)` observations and `r ncol(instacart)` variables. A few key variables in the dataset include an 'order_id', 'user_id', which are the same, indicating that these are variables specific to each instacart customer. Other important variables include 'product_id' and 'product_name', which give more details about the products purchased in their instacart order.

* A user with user_id 112108 placed an order (with order_id = 1) on the fourth day of the week (order_dow = 4) at 10 am (order_hour_of_day = 10). 9 days have passed since their previous order. In this order, they added "Bulgarian Yogurt" to their cart first (add_to_cart_order = 1), which is found in the "yogurt" aisle of the "dairy eggs" department. This was their fourth instacart order (order_number = 4).

* A user with user_id 17989 had `r nrow(user_17989)` items in their order, with the most popular department they shopped in being the department_id 4, which corresponds to fresh vegetables.

> How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle_count = 
  instacart |>  
  distinct(aisle_id) |> 
  nrow()

aisle_top3 = 
  instacart |> 
  group_by(aisle_id) |> 
  summarize(n_obs = n()) |> arrange(desc(n_obs)) |> slice_head(n = 3) 
```

There are `r aisle_count` total aisles. The top 3 aisles are in the table below!
 
 Aisle Number | Count
------------- | -------------
           83 | 150609
           24 | 150473 
          123 | 78493


> Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r, message=FALSE}
instacart |> 
  group_by(aisle_id, department) |> 
  summarize(n_obs = n()) |> 
  filter(n_obs > 10000)|> 
  ggplot( aes(x = as.factor(aisle_id), y = n_obs)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Items Ordered by Aisle & Department", x = "Aisle Number", y = "Number of Items") +
  theme_minimal() +
  facet_wrap(~ department, scales = "free_x") 
```

> Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
popular_items_tbl = 
  instacart |> janitor::clean_names() |> 
  select(aisle, product_name) |> 
  group_by(aisle,product_name) |> 
  summarize(n_obs = n()) |> filter(aisle %in% c("baking ingredients","dog food care", "packaged vegetables fruits")) |> arrange(desc(n_obs))
 



column_freq= instacart %>%
  count(aisle, sort = TRUE)
 
```



> Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


