---
title: "p8105_hw3_bac2214"
author: "Brianna Carnagie"
date: "2023-10-08"
output: github_document
---
### Question 1
```{r,echo=FALSE, message=FALSE}
library(p8105.datasets)
library(tidyverse)
library(ggplot2)
library(viridis)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```
#### Loading the data
```{r}
data("instacart") 
```
#### Exploring the data
```{r}
instacart_reordered = 
  instacart |> 
  select(order_id, user_id, order_number, everything())

user_112108 = 
  instacart |>  
  filter(user_id == 112108)

user_17989 = 
  instacart |>  
  filter(user_id == 17989) 

user_17989_dept_frq = 
  user_17989 |> 
  count(department_id)
```

The dataset is a tibble that has `r nrow(instacart)` observations and `r ncol(instacart)` variables. A few key variables in the dataset include an 'order_id', 'user_id', which are the same, indicating that these are variables specific to each instacart customer. Other important variables include 'product_id' and 'product_name', which give more details about the products purchased in their instacart order.

* A user with user_id 112108 placed an order (with order_id = 1) on the fourth day of the week (order_dow = 4) at 10 am (order_hour_of_day = 10). 9 days have passed since their previous order. In this order, they added "Bulgarian Yogurt" to their cart first (add_to_cart_order = 1), which is found in the "yogurt" aisle of the "dairy eggs" department. This was their fourth instacart order (order_number = 4).

* A user with user_id 17989 had `r nrow(user_17989)` items in their order, with the most popular department they shopped in being the department_id 4, which corresponds to fresh vegetables.

> How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle_count = 
  instacart |>  
  distinct(aisle_id) |> 
  nrow()

aisle_top3 = 
  instacart |> 
  group_by(aisle_id) |> 
  summarize(n_obs = n()) |> arrange(desc(n_obs)) |> slice_head(n = 3) 
```

There are `r aisle_count` total aisles. The top 3 aisles are in the table below!
 
 Aisle Number | Count
------------- | -------------
           83 | 150609
           24 | 150473 
          123 | 78493


> Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r, message=FALSE}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

> Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r, message=FALSE}
  instacart |> janitor::clean_names() |> 
  select(aisle, product_name) |> 
  group_by(aisle,product_name) |> 
  summarize(n_obs = n()) |> filter(aisle %in% c("baking ingredients","dog food care", "packaged vegetables fruits")) |> arrange(desc(n_obs)) |> slice_head(n = 3) |>  knitr::kable(digits = 4)
```

> Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |> janitor::clean_names() |> 
  select(order_dow, order_hour_of_day, product_name) |> 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

### Question 2

#### Loading the data
```{r}
data("brfss_smart2010")
```

#### Cleaning the dataset 
```{r}
brfss_df = brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health", response %in% c("Poor","Fair", "Good","Excellent")) |> 
  rename(crude_prevalence_percent = data_value) |> 
  select(year:sample_size, crude_prevalence_percent,everything(), -data_value_type, -data_value_unit) |> 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Excellent")))
```


> In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
states_2002 = brfss_df |> 
  filter(year == 2002) |> 
  group_by(locationabbr) |>      
  summarise(n_obs = n_distinct(locationdesc)) |>   
  filter(n_obs >= 7) |>      
  pull(locationabbr)   
  
states_2010 = brfss_df |> 
  filter(year == 2010) |> 
  group_by(locationabbr) |>      
  summarise(n_obs = n_distinct(locationdesc)) |>   
  filter(n_obs >= 7) |>      
  pull(locationabbr) 
```

The states that were observed at 7 or more locations in 2002 were `r states_2002`.
The states that were observed at 7 or more locations in 2010 were `r states_2010`.


> Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent_df = brfss_df |> 
  filter(response == "Excellent") |> 
  group_by(year, locationabbr) |> 
  summarise(avg_value = mean(crude_prevalence_percent, na.rm = TRUE)) |> 
  ungroup()

ggplot(excellent_df, aes(x = year, y = avg_value, group = locationabbr)) +
  geom_line(aes(color = locationabbr), alpha = 0.5) +
  scale_color_viridis(discrete = TRUE, guide = "none")  +
  labs(title = "Average 'Excellent' Response Value Over Time by State",
       x = "Year",
       y = "Crude Prevalence",
       color = "State")
```


> Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
two_panel_df = brfss_df |> 
 filter(year %in% c(2006, 2010)) |> 
  group_by(year, locationabbr) 

  
ggplot(two_panel_df, aes(x = response, y = crude_prevalence_percent)) +
  geom_boxplot() +
  facet_wrap(~ year, scales = "free_x") +
  labs(title = "Distribution of Data Value for Responses in NY State",
       x = "Response",
       y = "Crude Prevalence Percent") 
```

### Question 3

#### Loading the data
```{r}
nhanes_accel = read_csv("data/nhanes_accel.csv")
nhanes_covar = read_csv("data/nhanes_covar.csv", skip = 4)
```

Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

#### Cleaning and merging the datasets 
```{r}
nhanes_accel_tidy = nhanes_accel |> 
  janitor::clean_names() |> 
  pivot_longer(
     min1:min1440,
    names_to = "minute",
    names_prefix = "min",
    values_to = "mims") 
 
nhanes_covar_tidy = nhanes_covar |> 
  janitor::clean_names() |> 
  mutate(
    sex = case_when(
      sex == 1 ~ "male",
      sex == 2 ~ "female"
    ),
    education = case_when(
      education == 1 ~ "Less than high school",
      education == 2 ~ "High school equivalent",
      education == 3 ~ "More than high school"))

nhanes_merge_df = full_join(nhanes_accel_tidy, nhanes_covar_tidy, by = "seqn") |> 
  janitor::clean_names() |> 
  filter(
    age >= 21, 
    if_all(c(seqn, minute, mims, sex, age, bmi, education), ~ !is.na(.))
  ) |> 
  mutate(education = factor(education, levels = c("Less than high school", "High school equivalent", "More than high school")))
```


> Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}

gender_df = nhanes_merge_df |> 
  group_by(sex, education) |> 
  summarize(n_obs = n_distinct(seqn)) 

gender_tbl = gender_df |>  
  knitr::kable(digits = 4)


gender_plot = ggplot(gender_df, aes(x = education, y = n_obs)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Number of Participants by Education Level",
       x = "Education Level",
       y = "Number of Participants") +
  facet_wrap(~sex, ncol = 2)
```
* Based on the data, more females in the study completed more then high school compared to males.
* However, more males completed the equivalent of high school compared to females.


> Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}

```


> Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}

```


